---
title: "Tidyverse"
subtitle: "In Class Activity"
format: html
execute:
  echo: false
---


## The Question

> Describe the daytime air temperatures at the Rice Rivers Center for February, 2014 by day of the week.


## The Data

For this topic we will use some example data from the [Rice Rivers Center](https://ricerivers.vcu.edu).  These data represent both atmospheric and water data collected from instrumentation on-site.  I have stored these data in a spreadsheet that is shared on Google Drive as a CSV file.

```{r}
library( tidyverse )
url <- "https://docs.google.com/spreadsheets/d/1Mk1YGH9LqjF7drJE-td1G_JkdADOU0eMlrP01WFBT8s/pub?gid=0&single=true&output=csv"
read_csv( url ) %>%
  mutate( Date = mdy_hms(DateTime) ) %>%        
  select( -DateTime ) %>% 
  mutate( Weekday = wday(Date, label=TRUE, abbr = FALSE ) ) -> rice 


summary( rice )

```

These data represent measurements taken every 15 minutes, 24 hours a day, 7 days a week, 365 days a year.  For brevity, this file contains measurements starting at `r rice$DateTime[1]` and ending at `r rev(rice$DateTime)[1]` (only `r nrow(rice)` records here...).

If you look at the summary of the data above, you will see several things, including:

- Date and time objects are `character`
- Some measurements are in Standard and some in Imperial with units in the same file include both °F and °C, as well as measurements in meters, feet, and inches.  In fact, there are duplication of data columns in different units (guess what kind of correlation they might have...)



Q: Mean Rainfall by Day of the Week

```{r}

library( knitr )
library( kableExtra )

rice %>%   # load the data
  select( Weekday, Rain_in) %>%                   # select weekday
  mutate( Rain = Rain_in * 2.54 ) %>%                 # convert to cm
  group_by( Weekday ) %>%
  summarize( `Mean Rain (cm)` = mean(Rain) ) %>%
  kable(digits = 4 ) %>%
  kable_paper( full_width = FALSE)
  
  
# summarize Mean Rainfall
```






```{r}


```


Min & Max Turbidity by Month as Kable 

```{r}
rice %>%
  mutate( Month = month( Date,label = TRUE, abbr = FALSE ) ) %>%
  group_by( Month ) %>% 
  summarize( Minimum = min( Turbidity_ntu, na.rm=TRUE  ), 
              Maximum = max( Turbidity_ntu, na.rm=TRUE  ) ) 
  
```




```{r}

rice  %>%                                    # get the data 
  select( RelHumidity, AirTempF )  %>%       # select Rel Hum & Temp
  filter( AirTempF > 32 )          %>%        # filter
# mutate
# summarize 

```






## Weekly High Tide

Find the highest tide for each week in the dataset


```{r}

```




? High and low Chaluga for Month in Day or Night 

Month    | Daytime  | High | Mean | Low 
---------|----------|------|------|-----
January  | Day      |      |      |   
         | Night    |      |      | 
February | Day      |      |      |   
         | Night    |      |      | 




