---
title: "Rice Center Atmospheric & Aquatic Conditions"
subtitle: "Key to the Homework"
author: "Key"
execute: 
  echo: false 
---

```{r}
#| echo: false 
#| message: false 
options( tidyverse.quiet = TRUE )
options(dplyr.summarise.inform = FALSE)
library( knitr )
library( kableExtra )
library( tidyverse )
library( lubridate )
url <- "https://docs.google.com/spreadsheets/d/1Mk1YGH9LqjF7drJE-td1G_JkdADOU0eMlrP01WFBT8s/pub?gid=0&single=true&output=csv"
read_csv(url) |> 
  mutate( Date = mdy_hms(DateTime, tz="EST") ) |>
  mutate( Month = month(Date, abbr=FALSE, label=TRUE) ) -> data 
```



## Methods


The data for this analysis has been retrieved from atmoshperic and aquatic sensors located at the [Rice Rivers Center](https://ricerivers.vcu.edu) in Charles City County, Virginia.  All data are collected automatically every 15 minutes and stored in an online repository as comma-separated data sets.  For this report, readings will be used dating from January 1, 2014 until March 27, 2014.   



### Atmospheric Conditions

Atmospheric measurements included *Air Temperature*, *Relative Humidity*, *Precipitation*, and *Barometric Pressure* and were taken from the monitoring station located on the bluff overlooking the James River in the vicinity of the Conference Center. Where necessary, all data were converted to SI units.  Minimum, mean, and maximum values were recorded for each of the first three months of the sampling period and presented in tabular format.

```{r}
#| echo: false 

data |> 
  mutate( AirTemp = AirTempF +32 * 9 / 5 ) |>
  mutate( Precipitation = Rain_in * 2.54 ) |>
  mutate( Pressure = BP_HG ) |>
  select( Date, Month, 
          Temperature = AirTemp,
          `Relative Humidity` = RelHumidity, 
          Precipitation,
          `Barometric Pressure` = Pressure ) -> atmospheric

atmospheric |> 
  group_by( Month ) |> 
  summarize( Minimum = min( Temperature ), 
             Mean = mean( Temperature ),
             Max = max( Temperature )  )  -> temp_table 

atmospheric |> 
  group_by( Month ) |> 
  summarize( Minimum = min( `Relative Humidity` ), 
             Mean = mean( `Relative Humidity` ),
             Max = max( `Relative Humidity` )  ) -> humidity_table

atmospheric |> 
  group_by( Month ) |> 
  summarize( Minimum = min( Precipitation ), 
             Mean = mean( Precipitation ),
             Max = max( Precipitation )  ) -> precip_table 


atmospheric |> 
  group_by( Month ) |> 
  summarize( Minimum = min( `Barometric Pressure` ), 
             Mean = mean( `Barometric Pressure` ),
             Max = max( `Barometric Pressure` )  ) -> pressure_table 

```






### Aquatic Conditions


Aquatic measurements included *Water Temperature*, *Salinity*, *pH*, and *Turbidity* and were attained from sensors attached to the research dock on the James River. Where necessary, all data were converted to SI units.  Minimum, mean, and maximum values were recorded for each of the first three months of the sampling period and presented in tabular format.

```{r}
data |> 
  select( Date, Month, 
          Temp = H2O_TempC,
          Salinity = Salinity_ppt,
          pH = PH,
          Turbidity = Turbidity_ntu) -> aquatic 

aquatic |> 
  group_by( Month ) |> 
  summarize( Minimum = min( Salinity, na.rm=TRUE  ),
             Mean = mean( Salinity, na.rm=TRUE  ),
             Max = max( Salinity, na.rm=TRUE  ) ) -> salinity_table
aquatic |> 
  group_by( Month ) |> 
  summarize( Minimum = min( pH, na.rm=TRUE  ),
             Mean = mean( pH, na.rm=TRUE  ),
             Max = max( pH, na.rm=TRUE  ) ) -> pH_table
aquatic |> 
  group_by( Month ) |> 
  summarize( Minimum = min( Turbidity, na.rm=TRUE  ),
             Mean = mean( Turbidity, na.rm=TRUE  ),
             Max = max( Turbidity, na.rm=TRUE  ) ) -> turb_table
aquatic |> 
  group_by( Month ) |> 
  summarize( Minimum = min( Temp, na.rm=TRUE ),
             Mean = mean( Temp, na.rm=TRUE ),
             Max = max( Temp, na.rm=TRUE ) ) -> wtemp_table
             
```


## Results


### Atmospheric Conditions

The mean and range of all atmospheric conditions for the first three months of 2014 are provided in the tables below.

```{r} 
temp_table |>
  kable( digits = 2, 
         caption = "Range of atmospheric temperatures (°C) measured on the bluff overlooking the James River at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()

```



```{r} 
humidity_table |>
  kable( digits = 2, 
         caption = "Range of relative atmospheric humidity (%) measured on the bluff overlooking the James River at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()
```



```{r} 
precip_table |>
  kable( digits = 2, 
         caption = "Precipitation (in cm) measured on the bluff overlooking the James River at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()
```



```{r} 
pressure_table |>
  kable( digits = 2, 
         caption = "Barometric Pressure (mmHG) measured on the bluff overlooking the James River at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()

```



### Aquatic Conditions

The mean and range of all aquatic conditions for the first three months of 2014 are provided in the tables below.  For some measurements in the month of January, equipment malfunctioned and there were no readings for some time periods.  Those measurements were removed form the analyses and not included in the estimates below.

```{r}
wtemp_table |>
  kable( digits = 2, 
         caption = "Range of water temperatures (°C) measured on instriments on the reserach dock at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()

```

```{r}
salinity_table |>
  kable( digits = 2, 
         caption = "Range of water salinity (ppt) measured on instriments on the reserach dock at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()

```

```{r}
pH_table |>
  kable( digits = 2, 
         caption = "Range of acidity (quantified as pH) measured on instriments on the reserach dock at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()

```

```{r}
turb_table |>
  kable( digits = 2, 
         caption = "Range of aquatic turbidity measured on instriments on the reserach dock at the Rice Rivers Center for the first three months of 2014.") |>
  kable_styling()

```
